---
title: 'Logistic Regression: Lab 3'
author: "Julianna Renzi"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(GGally) # for pairs plots
library(broom) # to get tidy model outputs in dataframes
library(jtools) # quick graphs of model outputs
```

## Data exploration 

```{r, message=FALSE}
# make a pairs plot with GGally
penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color = species))

```

```{r}
# two species that overlap (harder to classify, gentoos are larger)
# note class of species is a factor (gentoo will stick around as a factor because of this)
levels(penguins$species)

# Create subset
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>%  # keep species in any part of that vector
  mutate(species = fct_drop(species)) # want to make it clear that there are only 2 levels (no Gentoo)--fct_drop drops factors where there are no observations. Now Adelie is first (0) and Chinstrap is second (1)



```

Dataviz of the two species:

```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)
```
See that there is some sexual dimorphism (males appear larger). If these were your data you would do more exploration here

## Binary logistic regression

- body mass
- flipper length
- penguin sex

```{r}
# make the binomial logistic regression (still a generalized linear model, just with a different link)
adelie_chinstrap_blr <- glm(species ~ body_mass_g + flipper_length_mm + sex,
                            data = adelie_chinstrap,
                            family = "binomial")
```

Take a look at the outcomes

```{r}
adelie_chinstrap_blr

summary(adelie_chinstrap_blr) # null hypothesis is that the coefficient = 0
# remember lefthand side is probability of chinstrap (non-zero level, "level 1")--so as bodymass increases, odds of a penguin being a chinstrap may be going down (although non-significant). When flipper length increases, odds that it's a chinstrap is going up. If a penguin is male, it is less likely to be a chinstrap penguin (that's a little weird maybe don't take that to heart)
```

Does it make sense that increasing flipper length results in increasing probability of being chinstrap?

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) + 
  geom_jitter(aes(color = sex)) # maybe chinstraps are a little larger--makes sense
```

Interpreting coefficients (as probabilities) using broom augment function

```{r}
# augment takes original data and then adds on column of probability based on model of an observation being chinstrap
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict = "response") # need to specify probabilities [response], not log odds
# fitted value is probability that the observation is a chinstrap
# doesn't always predict correctly 
# would want to make a matrix of false positives, true positives, etc. and see how well it did

```

Plot fitted probabilities:

```{r}
ggplot(data = blr_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = sex, pch = species)) + 
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = "Flipper length (mm)",
       y = "Probability of outcome Chinstrap")
# flipper length seems most significant
# could also add a layer with actual data points (zeros for Adelies and their flipper lengths, then 1's for Chinstraps and their flipper lengths) 
```

Effect plot from jtools package for quick overviews of what model looks like based on different variables

```{r}
effect_plot(adelie_chinstrap_blr,
            pred = flipper_length_mm,
            y.label = "Probability of Chinstrap",
            interval = TRUE) # this is actually the model fit

# body mass vs. being a chinstrap
effect_plot(adelie_chinstrap_blr,
            pred = body_mass_g,
            y.label = "Probability of Chinstrap",
            interval = TRUE) # this is actually the model fit
# this is less customizable, though, so better to be in ggplot if not just looking for a quick overview

```

Use predict function to ask about data not existing in one's dataset. For instance, what's the probability that a female penguin weighing 3410 pounds with a flipper length of 192 is  a chinstrap?

```{r}
ex_1 <- predict(adelie_chinstrap_blr,
                data.frame(sex = "female",
                           body_mass_g = 3410,
                           flipper_length_mm = 192),
                type = "response") # response is probability! Otherwise it's lod-odds

ex_1 # probability it's a chinstrap

```
